---
title: "Migration_Log_Reg"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load("logreg.RData")
```

## Logistic Regression using all data points
### Variables: Maximum SIC w/in 30 km radius, Reproductive Status, Distance to Land, 
### Don't bother with this - should distill data to one point per day

```{r, echo = FALSE}
fit_d2l <- glm(leave.ice ~ dist2land, data = logreg, family = binomial())
summary(fit_d2l)

fit_max <- glm(leave.ice ~ SIC_30m_max, data = logreg, family = binomial())
summary(fit_max)

fit_repro <- glm(leave.ice ~ repro, data = logreg, family = binomial())
summary(fit_repro)

fit_repro_max <- glm(leave.ice ~ repro + SIC_30m_max, data = logreg, family = binomial())

fit_d2l_max <- glm(leave.ice ~ dist2land + SIC_30m_max, data = logreg, family = binomial())

fit_d2l_max_repro <- glm(leave.ice ~ repro + SIC_30m_max + dist2land, data = logreg, family = binomial())

aicc <- AICc(fit_max, fit_d2l, fit_repro, fit_repro_max, fit_d2l_max, fit_d2l_max_repro)

# Learn how to call functions from another script

create_AICc_table(aicc) # From MyFunctions.R
```

# Results are cleaner when d2l is not included as a single model. Maybe d2l is absorbing too much variation. 

```{r, echo=FALSE}
fit_d2l <- glm(leave.ice ~ dist2land, data = loc1, family = binomial())
summary(fit_d2l)

fit_max <- glm(leave.ice ~ SIC_30m_max, data = loc1, family = binomial())
summary(fit_max)

fit_repro <- glm(leave.ice ~ repro, data = loc1, family = binomial())
summary(fit_repro)

fit_repro_max <- glm(leave.ice ~ repro + SIC_30m_max, data = loc1, family = binomial())

fit_d2l_max <- glm(leave.ice ~ dist2land + SIC_30m_max, data = loc1, family = binomial())

fit_d2l_max_repro <- glm(leave.ice ~ repro + SIC_30m_max + dist2land, data = loc1, family = binomial())

aicc <- AICc(fit_max, fit_d2l, fit_repro,fit_repro_max, fit_d2l_max, fit_d2l_max_repro)

# Learn how to call functions from another script

create_AICc_table(aicc) # From MyFunctions.R
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
